---
title: "Boruta feature selection using xgBoost with SHAP analysis"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
  encoding=encoding,
  output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(tidyr)
library(caret)
library(anytime)
library(e1071)
library(glmnet)

library(xgboost)
library(DiagrammeR)
library(ggplot2)
library(hrbrthemes)
library(viridis)
library(ggrepel)
library(SHAPforxgboost)
library(forcats)
library(Boruta)
library(reshape2)

setwd("G:/My Drive/WorkingDesk/GitHub page/BorutaFeatureSelectionWithShapAnalysis")

load("scaled.train.test.RData")
xgb.mod=readRDS("xgb_scaled.mod")
xgb.boruta=readRDS("xgb_scaled.boruta")
```

Assuming a tunned xgBoost algorithm is already fitted to a training data set, (e.g., [look at my own implementation](https://amirali-n.github.io/ExtremeGradientBoosting/)) the next step is to identify feature importances. Although, feature importances can be evalutated directly from the boosted trees, these importances have been shown to be local and inconsistent; see [Scott Lundberg et. al. (2019)](https://arxiv.org/abs/1905.04610).

Scott Lundberg's paper proposed a new method to interpret the result of machine learning algorithms, particularyly boosting methods, that produce feature importance scores. They show that feature importance, although averaged over boosted trees, may be inconsistent for different observations. To overcome this inconsistency, they propose a SHAP score inspired by [Shapley values](http://www.library.fa.ru/files/Roth2.pdf#page=39) which combines different explanation models, e.g., LIME, with Shapley values. The result is a global feature importance score that is consistent across different test sets.

However, other than to arbitrarily select an importance threshold beyond which features are considered unimportant, SHAP analysis does not offer an algorithmic way to filter a large feature set to a limited set of important features.

To that end, a selection wrapper algorithm known as [Boruta](https://pdfs.semanticscholar.org/85a8/b1d9c52f9f795fda7e12376e751526953f38.pdf%3E) is proposed that iterates over an extended set of features and judges their importance. In each iteration of the algorithm, a number of features are copied by shuffling their values. This is used to fit another learner and to re-evaluate feature importances. If importance of an original feature is significantly greater than its shuffled copy, that features is deemed important.

```{r boruta, eval=FALSE}
library(Boruta)
xgb.boruta=Boruta(train.df,
                  y=as.numeric(as.factor(label))-1,
                  maxRuns=100, 
                  doTrace=2,
                  holdHistory=TRUE,
                  getImp=getImpXgboost,
                  max.depth=xgb.train$bestTune$max_depth, 
                  eta=xgb.train$bestTune$eta, 
                  nthread=4, 
                  min_child_weight=xgb.train$bestTune$min_child_weight,
                  scale_pos_weight=sumwneg/sumwpos, 
                  eval_metric="auc", 
                  eval_metric="rmse", 
                  eval_metric="logloss",
                  gamma=xgb.train$bestTune$gamma,
                  nrounds=xgb.crv$best_iteration, 
                  objective="binary:logistic",
                  tree_method="hist",
                  lambda=0,
                  alpha=0)
```

Here, the <code>xgb.train</code> stores the result of a cross-validated grid search to tune xgBoost hyperparameter; see **Boruta_xgBoost_SHAP.R** in [this repository](https://github.com/AmirAli-N/BorutaFeatureSelectionWithShapAnalysis). <code>xgb.cv</code> stores the result of 500 iterations of xgBoost with optimized paramters to determine the best number of iterations.

After comparing feature importances, Boruta makes a decision about the importance of a variable. If the importance of the shuffled copy is greater than or equal to the importance of the original feature, the decision is to <code>reject</code> that feature. If the original feature's importance is significantly greater than its shuffled copy, then the decision is to <code>confirm</code> that feature. Otherwise, Boruta is <code>tentative</code> about the true feature importance and cannot make a confirmation or rejection decision. Below, the result of 100 iterations of the Boruta algorithm is extracted. We see that only 6 features are confirmed and Boruta is not sure about the other 4 features.

```{r boruta_dec, echo=TRUE, eval=TRUE}
#extract Boruta's decision
boruta_dec=attStats(xgb.boruta)
boruta_dec[boruta_dec$decision!="Rejected",]
```

For visualization purposes, here, the history of Boruta importance evaluation is extracted to generate a box and whiskers plot.

```{r boruta_history, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#get the names of each feature
imp_features=row.names(boruta_dec)[which(boruta_dec$decision!="Rejected")]
#get feature importance history
boruta.imp.df=as.data.frame(xgb.brouta$ImpHistory)
#keep only confirmed and tentative features
boruta.imp.df=boruta.imp.df[,names(boruta.imp.df)%in%imp_features]
#transform the data to a data frame with two columns: feature and importance value
boruta.imp.df=melt(boruta.imp.df)
#create a data frame by adding the decision for each feature as well
boruta.imp.df=cbind.data.frame(boruta.imp.df, 
                               decision=boruta.df$decision[match(boruta.imp.df$variable, 
                                                                 row.names(boruta.df))])
#reorder features data frame by the importance median value
feature_order=with(boruta.imp.df, reorder(variable, value, median, order = TRUE))
boruta.imp.df$variable=factor(boruta.imp.df$variable, levels = levels(feature_order))
```

```{r boruta_plot, echo=FALSE, eval=TRUE, warning=FALSE, fig.align="center", fig.height=7, fig.width=10}
boruta.feat=c("Lane closure", "Work length", "Collision density", 
              "Truck AADT", "Closure length", "ADT", "Peak AADT",
              "AADT", "Work duration", "Closure coverage")
ggplot(data = boruta.imp.df, aes(x=variable, y=value, fill=decision))+
  geom_boxplot()+
  coord_flip()+
  theme_ipsum(axis_title_just = "center")+
  theme(plot.title = element_blank(),
        axis.text.x = element_text(hjust = 0.5, 
                                   size=18, 
                                   family = "Century Gothic", 
                                   color = "black"),
        axis.title.x = element_text(size = 18, 
                                    family = "Century Gothic", 
                                    color = "black",
                                    margin = margin(15, 0, 0, 0)),
        axis.text.y = element_text(size=18, 
                                   family = "Century Gothic", 
                                   color = "black"),
        axis.title.y = element_text(size=18, 
                                    family = "Century Gothic", 
                                    color = "black",
                                    margin = margin(0, 15, 0, 0)),
        axis.line.x = element_line(size=1.2),
        axis.line.y = element_line(size=1.2),
        legend.title = element_text(size=18, 
                                    family = "Century Gothic", 
                                    color = "black"),
        legend.text = element_text(size=18, 
                                   family = "Century Gothic", 
                                   color = "black"),
        legend.position = "top")+
  scale_x_discrete(labels=rev(boruta.feat))+
  xlab("Features")+
  ylab("Boruta Importance")
```

The feature importances are plotted using ggplot <code>geom_boxplot</code>. They show that lane closure, work length, and collision density are the top three important features. Boruta is very effective in reducing the number of features from more than 700 to just 10. However, these importances may not be consistent with respect to the test set. SHAP values take each data point into consideration when evaluating the importance of a feature. The below code snippet, takes the outcome of an xgBoost fitted model in <code>xgb.mod</code> and evaluates its SHAP values. 

```{r shap_values, echo=TRUE, eval=FALSE}
library(SHAPforxgboost)
shap_values=shap.values(xgb_model = xgb.mod, X_train = dtest)
shap_values$mean_shap_score[1:10]
```
```{r shap_score, echo=FALSE, eval=TRUE}
shap_values$mean_shap_score[1:10]
```

It can be seen the the top 10 most important features of the SHAP tree explainer do not agree with the Boruta algorithm. Although, some features are important according to both methods, SHAP values consider new features such as <code>route 219</code>, <code>activity D41000</code>, or <code>county SB to SLO</code> as important in predicting the class of the test set data point. Furthermore, the order of importance for those features which re deemed as importance by both methods is not consistent. The code snippet below limits the shap score and test set to the features selected by the Boruta method and plots SHAP values for each data point in test set diluted by a factor of 20. 

```{r shap_boruta, echo=TRUE, eval=FALSE}
shap.feat=unique(boruta.imp.df$variable)
X=data.frame(as.matrix(dtest))
X=X[, names(X) %in% shap.feat]

#removing outliers for better visualization
inliers=which(X$work_duration<=0.2 & X$work_duration>=0 &
              X$work_length<=0.4 & X$work_length>=0 &
              X$road_adt<=0.8 & X$road_adt>=0 &
              X$peak_aadt<=0.3 & X$peak_aadt>=0 &
              X$aadt<=0.8 & X$aadt>=0 &
              X$truck_aadt<=0.4 & X$truck_aadt>=0 &
              X$collision_density11_12<=0.6 & X$collision_density11_12>=0 &
              X$closure_coverage<=1 & X$closure_coverage>=0 &
              X$closure_length<=0.2 & X$closure_length>=0)

#shap summary for boruta features
shap_score_filtered=data.frame(shap_values$shap_score)
shap_score_filtered=shap_score_filtered[inliers, names(shap_score_filtered) %in% shap.feat]

#plot
shap.plot.summary.wrap2(shap_score = shap_score_filtered, 
                        X = X[inliers,],
                        dilute=20)
```

```{r shap_plot, echo=FALSE, eval=TRUE, fig.align="center", fig.height=7, fig.width=10, warning=FALSE, message=FALSE}
pp=shap.plot.summary.wrap2(shap_score = shap_score_filtered, 
                        X = X[inliers,],
                        dilute=20)
shap.names=c("Truck AADT", "Closure length", "AADT", "Work length",
             "Closure coverage", "Collision density", "Work duration",
             "ADT", "Lane closure", "Peak AADT")
pp+theme_ipsum(axis_title_just = "center")+
  theme(plot.title = element_blank(),
        axis.text.x = element_text(hjust = 0.5, 
                                   size=18, 
                                   family = "Century Gothic", 
                                   color = "black"),
        axis.title.x = element_text(size = 18, 
                                    family = "Century Gothic", 
                                    color = "black",
                                    margin = margin(15, 0, 0, 0)),
        axis.text.y = element_text(size=18, 
                                   family = "Century Gothic", 
                                   color = "black"),
        axis.title.y = element_text(size=18, 
                                    family = "Century Gothic", 
                                    color = "black",
                                    margin = margin(0, 15, 0, 0)),
        axis.line.x = element_line(size=1.2),
        axis.line.y = element_line(size=1.2),
        legend.title = element_text(size=18, 
                                    family = "Century Gothic", 
                                    color = "black"),
        legend.text = element_text(size=18, 
                                   family = "Century Gothic", 
                                   color = "black"),
        legend.position = "right")+
  scale_x_discrete(labels=rev(shap.names), limits=rev(levels(pp$data$variable)))+
  xlab("Features")+
  ylab("Shapley Values")+
  guides(colour=guide_colorbar(title = "Feature Value",
                               title.position = "top",
                               label.position = "left",
                               barwidth = 0.2,
                               barheight = 30))
```

The SHAP summary plots shows that considering only the features selected by Boruta does not resolve the difference between SHAP values and Boruta importances. The resulting order of feautre importances is completely different in SHAP tree explainer method. However, assuming features selected by Boruta are actually the most important features, the SHAP summary plot allows us to see how each feature contributes to the classification process. For example, consider the case of <code>Lane closure</code> feature. The shop summary plot shows that higher values of Lane closure (i.e., Lane closure = 1) increase the probability of the positive class (which in this case denotes the probability of collision with roadside work zones). This analysis could be usefull if we had more confidence in the selected features and their importances.

To reconcile Boruta and SHAP analysis, a combination of these methods may be the solution. An algorithm that copies the features and shuffles their values, but evaluates the importance of the original and its copy using Shapley values, and tests whether original importance of a feature is significantly greater that its shuffled copy. This can be done by a library in python known as [BorutaShap](https://pypi.org/project/BorutaShap/). Here is my implementation of the algorithm in python.